Visual Diagnostics for Constrained Optimisation with Application to Guided Tours

### 2021-01-21 paper/zhang-cook-laa-langrene-menendez.Rmd
more work on noisy index simulation, WIP

```{r kol-better, fig.height = 6, fig.cap="One-D projection of \\code{norm\\_kol} index on \\code{boa6} data optimised by \\code{search\\_better} with 20 randomly generated seeds. Each point represents a basis in the original 6D space, reduced to 2D by PCA. The grey points are random bases generated from 6-D hollow sphere and the two yellow points represent the local maximum corresponds to basis when `V2` is found and the global maximum where `V7` is found. The points produced by the search algorithm is colored by whether the global or local maximum is found with the interpolated bases highlighted as a path and the final basis as an amplified point."}
```

```{r kol-random, fig.height = 6, fig.cap = "One-D projection of \\code{norm\\_kol} index on \\code{boa6} data optimised by \\code{search\\_better\\_random} with the same seeds. \\code{search\\_better\\_random} has a probabilistic acceptance implementation that would also accept a basis with lower index value. This design allows the optimiser to jump out of the local maximum and hence more instances find the global maximum."}
```

```{r kol-random-tuned, fig.height= 6, fig.cap = "One-D projection of \\code{norm\\_kol} index on \\code{boa6} data optimised by \\code{search\\_better\\_random} with an larger searching neighbourhood of 0.7. Further tuning of the parameter has been taken place to adjust the neighbourhood parameter `alpha` from 0.5 to 0.7. "}
```

The second experiment with the noisy index is to understand how the optimisers perform when a local maximum is present. The dataset used is \code{boa6} where `x2` and `x7` are informative. The two theoretical best bases are [`r matrix(c(0, 1, 0, 0, 0, 0), nrow = 6)`] and [`r matrix(c(0, 0, 1, 0, 0, 0), nrow = 6)`] with index value `r index <- tourr::norm_kol(nrow(boa6)); proj <- as.matrix(boa6) %*% matrix(c(0, 1, 0, 0, 0, 0), nrow = 6); index(proj)` and `r index <- tourr::norm_kol(nrow(boa6)); proj <- as.matrix(boa6) %*% matrix(c(0, 0, 1, 0, 0, 0), nrow = 6); index(proj)`, respectively. Hence, the global maximum happens when variable `x7` is found. Simulation is done with 20 randomly generated seeds on two optimisers: `search_better` (Figure \ref{fig:kol-better}) and `search_better_random` (Figure \ref{fig:kol-random} and Figure \ref{fig:kol-random-tuned}). Compared to Figure \ref{fig:kol-better} and \ref{fig:kol-random}, Figure \ref{fig:kol-random-tuned}further increases the neighbourhood parameter `alpha` from 0.5 to 0.7 to enlarge the search space.  The data object is collected for each simulation and computation is done to reduce the bases from its original 6D to a 2D points by principal component analysis so as to view the search paths.

In Figure \ref{fig:kol-better}, the starting point largely affects whether `search_better` will find the global maximum. The three seeds on row three: seed 9145, seed 2511, and seed 9209, and three on row four: seed 2888, seed 9334, and seed 9819 easily find `V7` because they are born in Rome. On the other end, the ones on the first two rows can only find `V2`, largely due to the fact that they do not even have chances to search near `V7`.

A comparison between `search_better` and `search_better_random` shows all four combinations where both, neither, or one of the two algorithms find the global maximum. Two iconic cases that shows how `search_better_random` improves `search_better` are seeds 4761 (position (1,4) in Figure \ref{fig:kol-better} and (2,2) in Figure \ref{fig:kol-random}) and seed 1842 ((2,2) in Figure \ref{fig:kol-better} and (4,5) in Figure \ref{fig:kol-random}). In both cases, `search_better_random` accepts inferior points at the initial iteration and this acceptance has later been proved to change the direction of the search and hence allows the optimiser to explore near the global optimum and in the end, find the global optimum. However, this change of search direction may sometimes have an adverse effect on `search_better_random` and change the search to the neighbourhood of `V2`. This happens on seed 9209 and seed 9982 (position (1,1) and (1,2) in Figure \ref{fig:kol-random}). The two cases that neither algorithm finds the global optimum happens on seed 6170 and seed 2757 (position (1,1) and (1,2) in Figure \ref{fig:kol-better} and position (1,5) and (1,3) in Figure \ref{fig:kol-random}). This is largely due to the fact that the searching neighbourhood is not large enough to allow the space near `V7` to be adequately explored.

In Figure \ref{fig:kol-random-tuned}, the neighbourhood parameter `alpha` is increased from a 0.5 default to 0.7 to solve the insufficiency of search neighbourhood. Remember that a candidate basis is generated via a linear combination of the current basis and a randomly generated basis on the surface of the sphere, an increase of `alpha` gives more weights on the randomly generated bases and hence allows a wider search. Seed 9982 ((2,4) in Figure \ref{fig:kol-random-tuned}) and seed 2463 ((3,2) in Figure \ref{fig:kol-random-tuned}) benefit from this and find the global maximum.  A noticeable feature of this enlarged search space is that the interpolation paths start to jump around the space but its usefulness is arguable. In seed 2888 (position (3,1)), seed 2986 (position (2,3)), seed 9209 (position (4,4)), just to name a few, this allows a near `V2` bases to switch to a near `V7` bases and as a result these cases find the global maximum. While in the case of seed 6746 (position (1,2)) and seed 2757 (position (1,5)), the jump happens from near `V7` basis to near `V2` basis, it does not come back and leaves those simulations in the local maximum.

>>>>

```{r kol-result, fig.height = 6, fig.cap="One-D projection of \\code{norm\\_kol} index on \\code{boa6} data optimised by \\code{search\\_better}, starting from 20 different seeds. "}
```

The second experiment with the noisy index is to understand how the optimisers perform when a local maximum is present. The dataset used is \code{boa6} where `x2` and `x7` are informative. The two theoretical best bases are [`r matrix(c(0, 1, 0, 0, 0, 0), nrow = 6)`] and [`r matrix(c(0, 0, 1, 0, 0, 0), nrow = 6)`] with index value `r index <- tourr::norm_kol(nrow(boa6)); proj <- as.matrix(boa6) %*% matrix(c(0, 1, 0, 0, 0, 0), nrow = 6); index(proj)` and `r index <- tourr::norm_kol(nrow(boa6)); proj <- as.matrix(boa6) %*% matrix(c(0, 0, 1, 0, 0, 0), nrow = 6); index(proj)`, respectively. Hence, the global maximum happens when variable `x7` is found.

<!-- Simulation is done with 20 randomly generated seeds on two optimisers: `search_better` (Figure \ref{fig:kol-better}) and `search_better_random` (Figure \ref{fig:kol-random} and Figure \ref{fig:kol-random-tuned}). Compared to Figure \ref{fig:kol-better} and \ref{fig:kol-random}, Figure \ref{fig:kol-random-tuned}further increases the neighbourhood parameter `alpha` from 0.5 to 0.7 to enlarge the search space.  The data object is collected for each simulation and computation is done to reduce the bases from its original 6D to a 2D points by principal component analysis so as to view the search paths. -->

### 2021-01-21 paper/zhang-cook-laa-langrene-menendez.tex
more work on noisy index simulation, WIP

{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-better-1}

@@

\caption[One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better} with 20 randomly generated seeds]{One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better} with 20 randomly generated seeds. Each point represents a basis in the original 6D space, reduced to 2D by PCA. The grey points are random bases generated from 6-D hollow sphere and the two yellow points represent the local maximum corresponds to basis when `V2` is found and the global maximum where `V7` is found. The points produced by the search algorithm is colored by whether the global or local maximum is found with the interpolated bases highlighted as a path and the final basis as an amplified point.}\label{fig:kol-better}
\end{figure}
\end{Schunk}

{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-random-1}

\caption{One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better\_random} with the same seeds. \code{search\_better\_random} has a probabilistic acceptance implementation that would also accept a basis with lower index value. This design allows the optimiser to jump out of the local maximum and hence more instances find the global maximum.}\label{fig:kol-random}
\end{figure}
\end{Schunk}

{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-random-tuned-1}

\caption[One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better\_random} with an larger searching neighbourhood of 0.7]{One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better\_random} with an larger searching neighbourhood of 0.7. Further tuning of the parameter has been taken place to adjust the neighbourhood parameter `alpha` from 0.5 to 0.7. }\label{fig:kol-random-tuned}
\end{figure}
\end{Schunk}

The second experiment with the noisy index is to understand how the
optimisers perform when a local maximum is present. The dataset used is
\code{boa6} where \texttt{x2} and \texttt{x7} are informative. The two
theoretical best bases are {[}0, 1, 0, 0, 0, 0{]} and {[}0, 0, 1, 0, 0,
0{]} with index value 0.176 and 0.235, respectively. Hence, the global
maximum happens when variable \texttt{x7} is found. Simulation is done
with 20 randomly generated seeds on two optimisers:
\texttt{search\_better} (Figure \ref{fig:kol-better}) and
\texttt{search\_better\_random} (Figure \ref{fig:kol-random} and Figure
\ref{fig:kol-random-tuned}). Compared to Figure \ref{fig:kol-better} and
\ref{fig:kol-random}, Figure \ref{fig:kol-random-tuned}further increases
the neighbourhood parameter \texttt{alpha} from 0.5 to 0.7 to enlarge
the search space. The data object is collected for each simulation and
computation is done to reduce the bases from its original 6D to a 2D
points by principal component analysis so as to view the search paths.

In Figure \ref{fig:kol-better}, the starting point largely affects
whether \texttt{search\_better} will find the global maximum. The three
seeds on row three: seed 9145, seed 2511, and seed 9209, and three on
row four: seed 2888, seed 9334, and seed 9819 easily find \texttt{V7}
because they are born in Rome. On the other end, the ones on the first
two rows can only find \texttt{V2}, largely due to the fact that they do
not even have chances to search near \texttt{V7}.

A comparison between \texttt{search\_better} and
\texttt{search\_better\_random} shows all four combinations where both,
neither, or one of the two algorithms find the global maximum. Two
iconic cases that shows how \texttt{search\_better\_random} improves
\texttt{search\_better} are seeds 4761 (position (1,4) in Figure
\ref{fig:kol-better} and (2,2) in Figure \ref{fig:kol-random}) and seed
1842 ((2,2) in Figure \ref{fig:kol-better} and (4,5) in Figure
\ref{fig:kol-random}). In both cases, \texttt{search\_better\_random}
accepts inferior points at the initial iteration and this acceptance has
later been proved to change the direction of the search and hence allows
the optimiser to explore near the global optimum and in the end, find
the global optimum. However, this change of search direction may
sometimes have an adverse effect on \texttt{search\_better\_random} and
change the search to the neighbourhood of \texttt{V2}. This happens on
seed 9209 and seed 9982 (position (1,1) and (1,2) in Figure
\ref{fig:kol-random}). The two cases that neither algorithm finds the
global optimum happens on seed 6170 and seed 2757 (position (1,1) and
(1,2) in Figure \ref{fig:kol-better} and position (1,5) and (1,3) in
Figure \ref{fig:kol-random}). This is largely due to the fact that the
searching neighbourhood is not large enough to allow the space near
\texttt{V7} to be adequately explored.

In Figure \ref{fig:kol-random-tuned}, the neighbourhood parameter
\texttt{alpha} is increased from a 0.5 default to 0.7 to solve the
insufficiency of search neighbourhood. Remember that a candidate basis
is generated via a linear combination of the current basis and a
randomly generated basis on the surface of the sphere, an increase of
\texttt{alpha} gives more weights on the randomly generated bases and
hence allows a wider search. Seed 9982 ((2,4) in Figure
\ref{fig:kol-random-tuned}) and seed 2463 ((3,2) in Figure
\ref{fig:kol-random-tuned}) benefit from this and find the global
maximum. A noticeable feature of this enlarged search space is that the
interpolation paths start to jump around the space but its usefulness is
arguable. In seed 2888 (position (3,1)), seed 2986 (position (2,3)),
seed 9209 (position (4,4)), just to name a few, this allows a near
\texttt{V2} bases to switch to a near \texttt{V7} bases and as a result
these cases find the global maximum. While in the case of seed 6746
(position (1,2)) and seed 2757 (position (1,5)), the jump happens from
near \texttt{V7} basis to near \texttt{V2} basis, it does not come back
and leaves those simulations in the local maximum.

>>>>

{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-result-1}

@@

\caption[One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better}, starting from 20 different seeds]{One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better}, starting from 20 different seeds. }\label{fig:kol-result}
\end{figure}
\end{Schunk}

The second experiment with the noisy index is to understand how the
optimisers perform when a local maximum is present. The dataset used is
\code{boa6} where \texttt{x2} and \texttt{x7} are informative. The two
theoretical best bases are {[}0, 1, 0, 0, 0, 0{]} and {[}0, 0, 1, 0, 0,
0{]} with index value 0.176 and 0.235, respectively. Hence, the global
maximum happens when variable \texttt{x7} is found.
